---
title: "TikTok Video View Summary"
format: html
---

# TikTok Video Analysis

This report provides a summary of the TikTok video view data.

## 1. Load the data

```{r}
library(tidyverse)

views <- read.csv("data/raw/video_view.csv")
```

## 2. Dataset overview

```{r}
nrow(views)

str(views)

head(views)

summary(views)

colSums(is.na(views))
```

## 3. Overall performance

```{r}
overview <- views %>%
  summarise(
    videos = n(),
    creators = n_distinct(creator_id),
    avg_length = mean(video_length_sec, na.rm = TRUE),
    avg_impressions = mean(impressions_n, na.rm = TRUE),
    avg_watched = mean(watched_n, na.rm = TRUE),
    avg_watch_rate = mean(watch_rate, na.rm = TRUE),
    avg_watch_share = mean(watched_n / impressions_n, na.rm = TRUE),
    avg_watch_seconds = mean(
      video_length_sec * watched_n / impressions_n,
      na.rm = TRUE
    )
  )

overview
```

## 4. Watch share

```{r}
views <- views %>%
  mutate(
    watched_pct = watched_n / impressions_n * 100
  )

head(views)
```

## 5. Most viewed video

```{r}
topvideo <- views %>%
  filter(impressions_n == max(impressions_n, na.rm = TRUE))

topvideo
```

## 6. Distribution of video length

```{r}
hist(
  views$video_length_sec,
  main = "Distribution of Video Length",
  xlab = "Video length (seconds)"
)
```

## 7. Videos with high watch rates

```{r}
goodvideos <- views %>%
  filter(watch_rate > 0.8)

nrow(goodvideos)
```

## 8. Top 10 videos by watch rate

```{r}
top_watch_rate <- views %>%
  select(
    video_id,
    creator_id,
    impressions_n,
    watched_n,
    watch_rate
  ) %>%
  arrange(desc(watch_rate)) %>%
  slice_head(n = 10)

top_watch_rate
```

## 9. Performance by creator

```{r}
creator_summary <- views %>%
  group_by(creator_id) %>%
  summarise(
    videos = n(),
    total_impressions = sum(impressions_n, na.rm = TRUE),
    average_watch_rate = mean(watch_rate, na.rm = TRUE)
  ) %>%
  arrange(desc(total_impressions))

creator_summary
```

## 10. Conclusion

The analysis provides an overview of video volume, reach, watch behaviour, and creator-level performance.